---
title: "R Notebook"
output: html_notebook
---

#Loading the data file and running the Dimension, Structure and Summary Stats
```{r}

library(readxl)


Data1 = read_excel(file.choose())
Data2 = Data1[, -c(1,5)]
str(Data2)
dim(Data2)
summary(Data2)


```

#Converting the respective dependent and independent variables to catagorical variables

```{r}
Data2$Personal_Loan = as.factor(Data2$Personal_Loan)
Data2$Securities_Account = as.factor(Data2$Securities_Account)
Data2$CD_Account = as.factor(Data2$CD_Account)
Data2$Online = as.factor(Data2$Online)
Data2$CreditCard = as.factor(Data2$CreditCard)
names(Data2)
str(Data2)
```


# Checking for null values in all the columns and doing a KNN impute for the null values in Family Members column
```{r}
colSums(is.na(Data2))
library(VIM)
Data3 = kNN(Data2, variable = "Family_members", k = 6)
summary(Data3)

```

# There are few negative values in experience which are replaced with zeros

```{r}
Data3$Experience[Data3$Experience<0] = 0


summary(Data3$Experience)
```


# Exploratory Data Analysis - Univariate

```{r}
boxplot(Data3$Age, Main  = "Age ditribution", xlab = "Age", horizontal = T) 

boxplot(Data3$Experience, Main  = "Experience ditribution", xlab = "Experience", horizontal = T)

boxplot(Data3$Income, Main  = "Income ditribution", xlab = "Income", horizontal = T)

boxplot(Data3$Family_members, Main  = "Family ditribution", xlab = "Family Members", horizontal = T)

boxplot(Data3$CCAvg, Main  = "Credit Card avg spend", xlab = "CCAvg", horizontal = T)

hist(Data3$CCAvg)

boxplot(Data3$Mortgage, Main  = "Mortgage", xlab = "Value of house mortgage", horizontal = T)

library(ggplot2)


ggplot(Data3) +
  aes(x = Data3$Personal_Loan) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

ggplot(Data3) +
  aes(x = Data3$Education) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

ggplot(Data3) +
  aes(x = Data3$Securities_Account) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

ggplot(Data3) +
  aes(x = Data3$CD_Account) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

ggplot(Data3) +
  aes(x = Data3$Online) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

ggplot(Data3) +
  aes(x = Data3$CreditCard) +
  geom_bar() + 
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))

```

# Exploratory Data Analysis - bivariate

```{r}
ggplot(data=Data3, aes(x=Personal_Loan, y=Age, fill=Personal_Loan))+geom_boxplot()

ggplot(data=Data3, aes(x=Personal_Loan, y=Experience, fill=Personal_Loan))+geom_boxplot()

ggplot(data=Data3, aes(x=Personal_Loan, y=Income, fill=Personal_Loan))+geom_boxplot()

ggplot(data=Data3, aes(x=Personal_Loan, y=Family_members, fill=Personal_Loan))+geom_boxplot()

ggplot(data=Data3, aes(x=Personal_Loan, y=CCAvg, fill=Personal_Loan))+geom_boxplot()



ggplot(Data3, aes(fill=Personal_Loan, x=Education, y=as.numeric(Education))) +
  geom_bar( stat="identity", position="fill")

ggplot(Data3, aes(fill=Personal_Loan, x=CD_Account, y=as.numeric(CD_Account))) +
  geom_bar( stat="identity", position="fill")

ggplot(Data3, aes(fill=Personal_Loan, x=Securities_Account, y=as.numeric(Securities_Account))) +
  geom_bar( stat="identity", position="fill")

ggplot(Data3, aes(fill=Personal_Loan, x=Online, y=as.numeric(Online))) +
  geom_bar( stat="identity", position="fill")

ggplot(Data3, aes(fill=Personal_Loan, x=CreditCard, y=as.numeric(CreditCard))) +
  geom_bar( stat="identity", position="fill")


```


#Splitting the data into train and test data in the ratio 8:2
```{r}
## split into training and test sets
library(caTools)

set.seed(123)
split = sample.split(Data3$Personal_Loan, SplitRatio = 0.80)
traindata = subset(Data3, split == TRUE)
testdata = subset(Data3, split == FALSE)

## Check if distribution of partition data is correct Testing dataset
prop.table(table(Data3$Personal_Loan))
prop.table(table(Data3$Personal_Loan))
```
Buling a fully grown tree initially and print it
```{r}
library(rpart)
library(rpart.plot)
set.seed(420)

tree_full = rpart(formula = Personal_Loan~., data = traindata, cp=-1, minsplit=2, minbucket=1)
rpart.plot(tree_full, cex=0.4)


boxcols <- c("palegreen3", "orange")[tree_full$frame$yval]

par(xpd=TRUE)
prp(tree_full, faclen = 0, cex = 0.3, extra = 1, box.col = boxcols)
```

```{r}
print(tree_full)

plotcp(tree_full)
```


Predict on the train and test data set

```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_full,traindata,type="class")
traindata$predict.score=predict(tree_full,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(Personal_Loan,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec
```

```{r}
## Predict using the CART model
testdata$predict.class=predict(tree_full,testdata,type="class")
testdata$predict.score=predict(tree_full,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(Personal_Loan,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```

#Finding the performance measures of the full tree on train and test data set
```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test')
df_fin

```

#remove predicted score and class before running other models
```{r}
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```

# Pruning using minbuckt and minsplit

```{r}

set.seed(123)
?rpart

tree_manual_prune=rpart(formula = Personal_Loan~., data = traindata, method="class",control = rpart.control(minsplit = 50,  minbucket = 15))



rpart.plot(tree_manual_prune, cex=0.8)
```

# Predict on the train data set and calculating the performance measures
```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_manual_prune,traindata,type="class")
traindata$predict.score=predict(tree_manual_prune,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(Personal_Loan,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec
```

#Predict on the test data set and calculating the performance measures

```{r}
## Predict using the CART model
testdata$predict.class=predict(tree_manual_prune,testdata,type="class")
testdata$predict.score=predict(tree_manual_prune,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(Personal_Loan,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```

```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_fin, df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test', 'MPruned tree_train', 'Mpruned tree_test')
df_fin
```

```{r}
# Remove predicted score and class before running other models
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```

```{r}

printcp(tree_full)
plotcp(tree_full)

bestcp=tree_full$cptable[which.min(tree_full$cptable[,"xerror"]),"CP"]
bestcp


ptree=prune(tree_full,cp=bestcp)
print(ptree)
rpart.plot(ptree, cex = 0.8)
```


```{r}
## Predict using the CART model
traindata$predict.class=predict(ptree,traindata,type="class")
traindata$predict.score=predict(ptree,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(Personal_Loan,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec

## Predict using the CART model
testdata$predict.class=predict(ptree,testdata,type="class")
testdata$predict.score=predict(ptree,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(Personal_Loan,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```

```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_fin, df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test', 'MPruned tree_train', 'MPruned tree_test', 'Best CP_tree_train', 'Best CP_ptree_test')
round(df_fin,2)
```

#Printing the best tree

```{r}
boxcols <- c("palegreen3", "orange")[tree_full$frame$yval]

par(xpd=TRUE)
prp(ptree, faclen = 0, cex = 0.6, extra = 1, box.col = boxcols)



```

```{r}
#Variable importance
library(caret)

ptree$variable.importance

df_cart=data.frame(round(ptree$variable.importance,2))
df_cart
write.csv(df_cart,"varimpcart.csv")
```

```{r}
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```

```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_manual_prune,traindata,type="class")
traindata$predict.score=predict(tree_manual_prune,traindata)

View(traindata)

```


```{r}
library(pROC)
roc_obj = roc(traindata$Personal_Loan, traindata$predict.score[,2])


plot(roc_obj, print.auc = T)
```




























